\chapter{Introduction}
Taking mRNA expression profiles from patients or cell lines can provide tremendous insight for both medical and research purposes.
Approximately 280,000 of these profiles are publicly available, but incomplete annotations bottleneck further analysis.
Automatically annotating these mRNA expression profiles, while not mission-critical, would ease this process.
A straightforward and comparably fast annotation method would be guilt by association (GBA) using the expression profiles themselves.
A significant issue with this approach originates from the fact that most of the samples involved are derived from a biopsy where various signals (e.g., biological pathways, non-cancer tissues, nonbiological effects) are intermingled in the final mRNA expression profiles. A dominating signal could significantly influence the annotation made for a given sample using GBA, producing a biased annotation. Independent component analysis (ICA) has been utilised to circumvent this issue.

The collection and processing of this transcriptional data results in the starting point for this project — a database containing expression levels associated with their genes for some 280,000 samples.
Before commencing this project, a subset of the available data was manually annotated by collecting annotation information from the corresponding publications.
This work placement project will aim to automate the annotation of mRNA expression profiles using this partially annotated data to train a machine learning algorithm.

\section{Aims}
This work placement will aim to train a machine-learning algorithm to annotate mRNA expression profiles correctly.
If time allows, a web application will be developed to make the model accessible to other parties.

\begin{itemize}
    \item Primary: Train a machine learning algorithm to annotate mRNA expression profiles correctly
    \begin{itemize}
        \item Data exploration
        \item Model selection \& training
        \item Model evaluation
    \end{itemize}
    \item Secondary: Develop a web application to annotate user's mRNA expression profiles using the trained model.
    \begin{itemize}
        \item Infrastructure
        \item Backend logic
        \item Frontend design
    \end{itemize}
\end{itemize}
